---
title: "SL_proj_doc"
author: "NC, AdT"
date: '2022-07-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Learning mod B project

#Libraries and Functions load_dataset

```{r}
library(lubridate)
library(rmarkdown)
library(knitr)
library(ggplot2)

#loading dataset. Dataset need to be in the same folder of the project.
load_dataset = function(data_folder, file){
  PATH = paste(getwd(), data_folder, file, sep="/")
}
```

Load dataset and remove col that are not useful for our purpose

```{r}
tripdata_2019 = read.csv(load_dataset("SL_dataset", "bluebikes_tripdata_2019.csv"))
tripdata_2019_r = subset(tripdata_2019, select = -c(usertype, gender, start.station.name, end.station.name))
colnames(tripdata_2019_r)
rm(tripdata_2019)

# Control the number of trip longer than 1H and shorter than 1m
nrow(tripdata_2019_r[tripdata_2019_r$tripduration>3600,])
nrow(tripdata_2019_r[tripdata_2019_r$tripduration<60,])

# Split timestamp in date - time Then split date in y/m/d
tripdata_2019_r = tidyr::separate(tripdata_2019_r, starttime, c("startdate", "starttime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, startdate, c("syear", "startmonth", "startday"), sep = "-")

tripdata_2019_r = tidyr::separate(tripdata_2019_r, stoptime, c("stopdate", "stoptime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, stopdate, c("stopyear", "stopmonth", "stopday"), sep = "-")

# Remove redundant informations
tripdata_2019_r = subset(tripdata_2019_r, select = -c(syear, stopyear))

```

In order to avoid future possible issue with our analysis, we removed all the trips that started in a month and finished in the next one (ex. start as 31/07 23:59 and end 01/08 00:30). This specific kind of trip amount to 0.035% of the total.

```{r}
between_month = nrow(tripdata_2019_r[tripdata_2019_r$startmonth != tripdata_2019_r$stopmonth,])
ratio = round((as.integer(between_month) / as.integer(nrow(tripdata_2019_r))) * 100, 3)
tripdata_2019_r = subset(tripdata_2019_r, select = -c(stopmonth, startmonth))
```

```{r}
# Convert time to sec [starting from midnight]
tripdata_2019_r$starttime = hms(tripdata_2019_r$starttime)
tripdata_2019_r$starttime = as.numeric(tripdata_2019_r$starttime)
tripdata_2019_r$starttime = round(tripdata_2019_r$starttime,0)

tripdata_2019_r$stoptime = hms(tripdata_2019_r$stoptime)
tripdata_2019_r$stoptime = as.numeric(tripdata_2019_r$stoptime)
tripdata_2019_r$stoptime = round(tripdata_2019_r$stoptime,0)
```

```{r}
par(mfrow=c(1,2))

freq_month_19 = barplot(months_19, 
     main = "Freq during months",
     xlab = "Months",
     ylab = "Freq",
     col = "light Blue")
text(x = freq_month_19, y = months_19 + 0.5, labels = months_19, cex=0.8)


mean_dur_19 = c()

for (val in 1:12){
  mean_dur_19[val] =  round(mean(tripdata_2019_r$tripduration[tripdata_2019_r$month == val])/60,2)
}

min_per_month_19 = barplot(mean_dur_19, 
     main = "Avg (min) per month",
     xlab = "Months",
     ylab = "Mins",
     col = "light green",
     names.arg=c("1","2","3","4","5", "6", "7", "8", "9", "10", "11", "12"))
text(x = min_per_month_19, y = mean_dur_19 + 0.5, labels = mean_dur_19, cex=0.7)
```

```{r}
tripdata_2019_r$age = tripdata_2019_r$year - tripdata_2019_r$birth.year
```

```{r}
cat("min age:", min(tripdata_2019_r$age))
cat("\nmax age:", max(tripdata_2019_r$age))
cat("\nusers >80y:", nrow(tripdata_2019_r[tripdata_2019_r$age > 80,]))
```
```{r}
nrow(tripdata_2019_r[tripdata_2019_r$age > 60 & tripdata_2019_r$age < 80,])
```



```{r}
hist(tripdata_2019_r$age, 
     breaks = 20,
     labels = TRUE)


dens = density(tripdata_2019_r$age)
plot(dens)
```



```{r}
boxplot(tripdata_2019_r$age)
```







Repeat the same exact procedure for the dataset of the 2020

```{r}

tripdata_2020 = read.csv(load_dataset("SL_dataset", "bluebikes_tripdata_2020.csv"))
tripdata_2020_r = subset(tripdata_2020, select = -c(usertype, gender, start.station.name, end.station.name))

tripdata_2020_r = tidyr::separate(tripdata_2020_r, starttime, c("startdate", "starttime"), sep = " ")
tripdata_2020_r = tidyr::separate(tripdata_2020_r, startdate, c("syear", "startmonth", "startday"), sep = "-")
tripdata_2020_r = tidyr::separate(tripdata_2020_r, stoptime, c("stopdate", "stoptime"), sep = " ")
tripdata_2020_r = tidyr::separate(tripdata_2020_r, stopdate, c("stopyear", "stopmonth", "stopday"), sep = "-")

tripdata_2020_r = subset(tripdata_2020_r, select = -c(syear, stopyear))

tripdata_2020_r$starttime = hms(tripdata_2020_r$starttime)
tripdata_2020_r$starttime = as.numeric(tripdata_2020_r$starttime)
tripdata_2020_r$starttime = round(tripdata_2020_r$starttime,0)

tripdata_2020_r$stoptime = hms(tripdata_2020_r$stoptime)
tripdata_2020_r$stoptime = as.numeric(tripdata_2020_r$stoptime)
tripdata_2020_r$stoptime = round(tripdata_2020_r$stoptime,0)

between_month = nrow(tripdata_2020_r[tripdata_2020_r$startmonth != tripdata_2020_r$stopmonth,])
print("% of trips between months in 2020", ratio = round((as.integer(between_month) / as.integer(nrow(tripdata_2020_r))) * 100, 3)) #0.039%

tripdata_2020_r = subset(tripdata_2020_r, select = -c(stopmonth, startmonth))

colnames(tripdata_2019)
colnames(tripdata_2020) #Remark: 2020 dataset has PostalCode, while 2019 no

#Remove old dataset to free memory
rm(tripdata_2019)
rm(tripdata_2020)


```

```{r}
tripdata_2020_r$birth.year[is.na(tripdata_2020_r$birth.year)] = ("not_dec")
```
