---
title: "SL_proj_doc"
author: "NC, AdT"
date: '2022-07-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Learning mod B project

#Libraries and Functions 
load_dataset: loading dataset. Dataset need to be in the same folder of the project.
```
library(lubridate)
library(rmarkdown)
library(knitr)

load_dataset = function(data_folder, file){
  PATH = paste(getwd(), data_folder, file, sep="/")
}
```

# 
Load dataset and remove col that are not useful for our purpose
```
tripdata_2019 = read.csv(load_dataset("SL_dataset", "bluebikes_tripdata_2019.csv"))
tripdata_2019_r = subset(tripdata_2019, select = -c(usertype, gender, start.station.name, end.station.name))
colnames(tripdata_2019_r)
```

# Control the number of trip longer than 1H and shorter than 1m
nrow(tripdata_2019_r[tripdata_2019_r$tripduration>3600,])
nrow(tripdata_2019_r[tripdata_2019_r$tripduration<60,])

# Split timestamp in date - time Then split date in y/m/d
tripdata_2019_r = tidyr::separate(tripdata_2019_r, starttime, c("startdate", "starttime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, startdate, c("syear", "startmonth", "startday"), sep = "-")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, stoptime, c("stopdate", "stoptime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, stopdate, c("stopyear", "stopmonth", "stopday"), sep = "-")

# Remove redundant informations
tripdata_2019_r = subset(tripdata_2019_r, select = -c(syear, stopyear))

```










When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```
