---
title: "SL_proj_doc"
author: "NC, AdT"
date: '2022-07-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Learning mod B project

#Libraries and Functions load_dataset

```{r}
library(lubridate)
library(rmarkdown)
library(knitr)

#loading dataset. Dataset need to be in the same folder of the project.
load_dataset = function(data_folder, file){
  PATH = paste(getwd(), data_folder, file, sep="/")
}
```

Load dataset and remove col that are not useful for our purpose

```{r}
tripdata_2019_r = read.csv(load_dataset("SL_dataset", "bluebikes_tripdata_2019.csv"))
colnames(tripdata_2019_r)
rm(tripdata_2019)

# Control the number of trip longer than 1H and shorter than 1m
nrow(tripdata_2019_r[tripdata_2019_r$tripduration>3600,])
nrow(tripdata_2019_r[tripdata_2019_r$tripduration<60,])

# Split timestamp in date - time Then split date in y/m/d
tripdata_2019_r = tidyr::separate(tripdata_2019_r, starttime, c("startdate", "starttime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, startdate, c("syear", "startmonth", "startday"), sep = "-")

tripdata_2019_r = tidyr::separate(tripdata_2019_r, stoptime, c("stopdate", "stoptime"), sep = " ")
tripdata_2019_r = tidyr::separate(tripdata_2019_r, stopdate, c("stopyear", "stopmonth", "stopday"), sep = "-")

# Remove redundant informations
tripdata_2019_r = subset(tripdata_2019_r, select = -c(syear, stopyear))

```

In order to avoid future possible issue with our analysis, we removed all the trips that started in a month and finished in the next one (ex. start as 31/07 23:59 and end 01/08 00:30). This specific kind of trip amount to 0.035% of the total.

```{r}
between_month = nrow(tripdata_2019_r[tripdata_2019_r$startmonth != tripdata_2019_r$stopmonth,])
ratio = round((as.integer(between_month) / as.integer(nrow(tripdata_2019_r))) * 100, 3)
tripdata_2019_r = subset(tripdata_2019_r, select = -c(stopmonth, startmonth))
```

```{r}
# Convert time to sec [starting from midnight]
tripdata_2019_r$starttime = hms(tripdata_2019_r$starttime)
tripdata_2019_r$starttime = as.numeric(tripdata_2019_r$starttime)
tripdata_2019_r$starttime = round(tripdata_2019_r$starttime,0)

tripdata_2019_r$stoptime = hms(tripdata_2019_r$stoptime)
tripdata_2019_r$stoptime = as.numeric(tripdata_2019_r$stoptime)
tripdata_2019_r$stoptime = round(tripdata_2019_r$stoptime,0)
```

```{r}
tripdata_2019_r$age = tripdata_2019_r$year - tripdata_2019_r$birth.year
```

```{r}
tripdata_2019_r = tripdata_2019_r[tripdata_2019_r$start.station.latitude > 1 & tripdata_2019_r$end.station.latitude >1,]
```

```{r}
#Gender, self-reported by member (Zero=unknown; 1=male; 2=female)
tripdata_2019_r$usertype[tripdata_2019_r$usertype == "Subscriber"] = 1
tripdata_2019_r$usertype[tripdata_2019_r$usertype == "Customer"] = 0
```

```{r}
write.csv(tripdata_2019_r, "C:/Users/Utente/Desktop/SL_project/tripdata_2019_r.csv", row.names = FALSE)
```







-----2020----------------------------------------------------------------------------------------------------------




Repeat the same exact procedure for the dataset of the 2020

```{r}
tripdata_2020_r = read.csv(load_dataset("SL_dataset", "bluebikes_tripdata_2020.csv"))
dec_2020 = read.csv(load_dataset("SL_dataset", "2020_12.csv"))

a = colnames(tripdata_2020_r)
b = colnames(dec_2020)
setdiff(a,b)

dec_2020$year = 2020
dec_2020$month = 12
dec_2020$birth.year = 0
dec_2020$gender = 0

tripdata_2020_r = rbind(tripdata_2020_r, dec_2020)


tripdata_2020_r = subset(tripdata_2020_r, select = -c(postal.code))
colnames(tripdata_2020_r)

# Control the number of trip longer than 1H and shorter than 1m
nrow(tripdata_2020_r[tripdata_2020_r$tripduration>3600,])
nrow(tripdata_2020_r[tripdata_2020_r$tripduration<60,])

# Split timestamp in date - time Then split date in y/m/d
tripdata_2020_r = tidyr::separate(tripdata_2020_r, starttime, c("startdate", "starttime"), sep = " ")
tripdata_2020_r = tidyr::separate(tripdata_2020_r, startdate, c("syear", "startmonth", "startday"), sep = "-")

tripdata_2020_r = tidyr::separate(tripdata_2020_r, stoptime, c("stopdate", "stoptime"), sep = " ")
tripdata_2020_r = tidyr::separate(tripdata_2020_r, stopdate, c("stopyear", "stopmonth", "stopday"), sep = "-")

# Remove redundant informations
tripdata_2020_r = subset(tripdata_2020_r, select = -c(syear, stopyear))

between_month = nrow(tripdata_2020_r[tripdata_2020_r$startmonth != tripdata_2020_r$stopmonth,])
ratio = round((as.integer(between_month) / as.integer(nrow(tripdata_2020_r))) * 100, 3)
tripdata_2020_r = subset(tripdata_2020_r, select = -c(stopmonth, startmonth))

# Convert time to sec [starting from midnight]
tripdata_2020_r$starttime = hms(tripdata_2020_r$starttime)
tripdata_2020_r$starttime = as.numeric(tripdata_2020_r$starttime)
tripdata_2020_r$starttime = round(tripdata_2020_r$starttime,0)

tripdata_2020_r$stoptime = hms(tripdata_2020_r$stoptime)
tripdata_2020_r$stoptime = as.numeric(tripdata_2020_r$stoptime)
tripdata_2020_r$stoptime = round(tripdata_2020_r$stoptime,0)

#Gender, self-reported by member (Zero=unknown; 1=male; 2=female)
tripdata_2020_r$gender[is.na(tripdata_2020_r$gender)] = 0
tripdata_2020_r$birth.year[is.na(tripdata_2020_r$birth.year)] = 0

tripdata_2020_r$age = tripdata_2020_r$year - tripdata_2020_r$birth.year
tripdata_2020_r$age[tripdata_2020_r$age == 2020] = 0
tripdata_2020_r = tripdata_2020_r[tripdata_2020_r$start.station.latitude > 1 & tripdata_2020_r$end.station.latitude >1,]

tripdata_2020_r$usertype[tripdata_2020_r$usertype == "Subscriber"] = 1
tripdata_2020_r$usertype[tripdata_2020_r$usertype == "Customer"] = 0

```

```{r}
write.csv(tripdata_2020_r, "C:/Users/Utente/Desktop/SL_project/tripdata_2020_r.csv", row.names = FALSE)
```



